---
title: "Fitanalyzer workflow example"
author: "Niraj Juneja, Charles de Lassence"
date: "March 9, 2016"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(fitcoach)
library(ggplot2)
```

## Example 1 : Data Loader- Getting data from Fitbit API

This part explains how to connect to the Fitbit API and get your data, using DataLoader. 


Step 1: You first need to make sure that you have [registered an app](https://dev.fitbit.com/) and set it as *Personal* in order to retrieve intraday data. You will need the following credentials in order to connect the API: App name (or *OAuth 2.0 Client ID*), Client Key and Client Secret. 

Step 2: We initialize a new DataLoader object, and connect to the API with Oauth2, using the credentials described above.

```{r eval = FALSE}
mydata <- DataLoader$new()
mydata$connect(appname = "cdlr",
               key = "227FWR",
               secret = "3089e3d1ac5dde1aa00b54a0c8661f42"
)
```

Step 3: We request the data and write it to JSON files using the `request` method. You need to specify the type of timeseries ('day' or 'intraday'), the list of activities ([full list here](https://dev.fitbit.com/docs/activity/)), the start and end dates, and the folder in which the JSON files will be written.

```{r eval = FALSE}
masterPath <- system.file("extdata", 
                          "daily-time-series", 
                          package = "fitcoach")

mydata$request(
    type = 'day', 
    activities = list("calories", "steps", "distance", "minutesVeryActive"), 
    start.date = "2016-01-01", 
    end.date = "2016-02-01", 
    path = masterPath
)
```

Once the JSON files have been created, they can be used for further analysis.


## Example 2 : Fit Analyzer - Daily File Analysis

Examples below demonstrate usage scenarios for FitAnalyzer

Step 1: We first need to point to a folder that contains the Json files for "daily" file analysis. Refer These files are created by DataLoader.R

We then create a new instance of Fitanalyzer passing in the folder and the goal that we want to optimize on. Goals can be the following a) calories b)steps c) distance d)floors 

The example below uses *steps* as the goal


```{r}
masterPath <- system.file("extdata", 
                          "daily-time-series", 
                          package = "fitcoach")

ana <- FitAnalyzer$new("steps")
```

Step 2: Next we get the data.frame ready for analysis. Note this data.frame is cleaned and augmented with additional data elements not present in the json file. eg: we augment weekday, weekend and mark rows that are valid

```{r}
timeseries.frame <- 
    ana$getAnalysisFrame(folder = masterPath, 
                         analysis.type = "daily")
head(timeseries.frame)
```


Step 3: next we find the most important variables that are enabling meeting the goals for the person. Note this call creates a glm model behind the scenes and ranks the variables based on the coefficients of the glm model. You can also get the glm fit object to do further analysis


```{r}
vars <- ana$findImportantVariables(tsDataFrame = timeseries.frame, 
                                   seed = 12345)
vars
```

Getting the fit object

```{r}
fit <- ana$getFit()
summary(fit)
```



```{r}
par(mfrow=c(2,2))
plot(fit)
```

Step 4: Next we can then plot the performance of the individual relative to the most important variables that are making a difference. 

```{r}
ana$showMostImportantCharts(tsDataFrame = timeseries.frame)
```


Step 5: We can also get the prediction on goal performance using the call below

```{r}
rows.test <- timeseries.frame[sample(1:191 , 1), ]
x <- createDependentVariableFrame(master = rows.test, goal = "steps")
res <- ana$predictGoal(x)
cat(paste("Prediction for the day" , "expected steps = ", round(res)))
```



*** 

## Example 3 :FitAnalyzer - Intra-day File Analysis
Examples below demonstrate usage scenarios for FitAnalyzer for **Intra day analysis**

*Step 1*: We first need to point to a folder that contains the Json files for *intraday* file analysis. Refer These files are created by DataLoader.R

We then create a new instance of Fitanalyzer passing in the folder and the goal that we want to optimize on. Goals can be the following a) calories b)steps c) distance d)floors 

The example below uses *calories* as the goal

```{r }
masterPath <-
    system.file("extdata", "intra-daily-timeseries", package = "fitcoach")
ana <- FitAnalyzer$new("calories")
```

Step 2: Next we get the data.frame ready for analysis. Note this data.frame is cleaned and augmented with additional data elements not present in the json file. eg: we augment cumulative sum during the day,weekday, weekend etc.

```{r}
intra <- ana$getAnalysisFrame(folder = masterPath, analysis.type = "intra.day")
head(intra)
```

Step 3: next we find the most important variables that are enabling meeting the goals for the person. Note:this call creates a **gbm** model behind the scenes and ranks the variables based on *relative.influence* call to gbm model. You can also get the gbm fit object to do further analysis

```{r}
vars <- ana$findImportantVariables(intra)
vars <- sort(vars, decreasing = TRUE)
vars
```
Plot of important variables below

```{r}
vars.frame <- data.frame(variables = names(vars), values = vars)
vars.frame$lnvalue <- log(vars.frame$values)
vars.frame <- vars.frame[1:7, ]
barplot(vars.frame$value, xlab = "variables", ylab = "relative importance", 
        names.arg = vars.frame$variables,
        cex.names = 0.65, cex.lab = 0.65, ylim = c(0.0, 0.1))
```

Summary of GBM model fit below

```{r}
fit <- ana$getFit()
summary(fit)
```

Step 4: Next we can then plot the performance of the individual relative to the most important variables that are making a difference. 
For the 4 most important variables, the average value for every 15 min of a day is plotted, along with the moving average (using `geom_smooth` from `ggplot2`), 

```{r}
ana$showMostImportantCharts(tsDataFrame = intra)
```

Step 5: We can also get the prediction on goal performance using the call below

```{r}
rows.test <- intra[sample(1:191 , 1), ] # take any random input for test
res <- ana$predictGoal(rows.test)
cat(paste("Prediction for the day" , "expected calories = ", round(res)))
```

